\textbf{zklora: Verifiable Multi-Party Inference for Evaluating Private LoRA Weights}

### 2025-01-12 paper/main.tex
SI to MPI, modified info flow a la distro

\title{\textbf{zklora: Verifiable Split Inference for Evaluating Private LoRA Weights}}
\author{Anonymous Authors}
\date{\today}

@@

\begin{abstract}
In decentralized computing environments, one party may wish to fine-tune a publicly available base model with Low-Rank Adaptation (LoRA) but keep those adapter weights private, while another party wants to evaluate performance on new data. 
We introduce \texttt{zklora}, a library designed to help the \emph{Base Model User} \emph{verifiably assess} the quality of a private LoRA module for a target task---for instance, to run a validation experiment without directly accessing the private weights. 
By adopting a split inference approach and generating \emph{verifiable computations}, \texttt{zklora} enables the \emph{Base Model User} to confirm that remote LoRA computations are correct, all without seeing the private weights themselves. 
We demonstrate the feasibility of \texttt{zklora} via benchmark experiments, showing that our approach remains computationally viable across various model sizes, with notably fast verification times.
\end{abstract}

\section{Introduction}
Large Language Models (LLMs) excel at a variety of tasks but come with significant computational and storage costs \cite{brown2020language, devlin2018bert}. 
Low-Rank Adaptation (LoRA) \cite{hu2021lora} addresses these costs by introducing a small number of trainable parameters into the model, drastically reducing memory overhead for fine-tuning \cite{ding2022delta}.

In this paper, we present \texttt{zklora}, a library that provides a \emph{verifiable computation} solution for the \emph{Base Model User} to \emph{evaluate} a private LoRA module. By orchestrating a split inference, \texttt{zklora} allows the Base Model User to run the base model locally while sending partial activations to the LoRA Owner. Crucially, the LoRA Owner returns not only updated activations but also a cryptographic \emph{proof} that the correct LoRA transformation was applied. In this way, the Base Model User can confirm the final outputs truly reflect the LoRA module’s capabilities. After verifying and measuring results on a validation dataset, the Base Model User can later choose how to adopt the LoRA at larger scale.

\section{Related Work}

\section{Methodology: zklora for Verifiable Evaluation}
Our aim is to enable the \emph{Base Model User} to run a \emph{validation experiment} on their local data, incorporating the private LoRA module from the \emph{LoRA Owner}. The Base Model User can measure task performance (e.g., cross-entropy loss, accuracy) to decide whether the LoRA meets performance requirements. Figure~\ref{fig:toy-mlp-lora} illustrates a simple scenario.

\begin{enumerate}
    \item \textbf{Base Model Forward Pass (Base Model User).}
    The \emph{Base Model User} runs layers without LoRA parameters locally, generating partial activations.

@@

\draw[arrow] (input) -- node[above, font=\scriptsize]{local multiply} (W1);
\draw[->, dashed] (W1.north) to[bend left=15] node[left, font=\scriptsize]{partial activations} (LoRA.south);
\draw[->, dashed] (LoRA.south) to[bend left=15] node[right, font=\scriptsize]{updated $\mathbf{z}'$} (W1.north);
\draw[arrow] (W1) -- node[above, font=\scriptsize]{$\mathbf{z}'$} (act);
\draw[arrow] (act) -- node[above, font=\scriptsize]{hidden} (W2);
\draw[arrow] (W2) -- (output);

\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black]above:\textbf{Base Model User}},
      inner sep=0.4cm
] (baseModelFit) {};

\end{tikzpicture}
\caption{A simplified 2-layer MLP with LoRA on the first layer. The Base Model User (blue box) runs most computations locally but calls the LoRA Owner (dashed box) to update partial activations. After verifying correctness, the Base Model User can measure final metrics to decide whether to adopt the private LoRA more extensively.}
\label{fig:toy-mlp-lora}
\end{figure}

After confirming performance with \texttt{zklora}, the Base Model User can move forward using the LoRA weights with confidence, knowing their underlying quality combined with the base model. The core objective with \texttt{zklora} is to \emph{verify} that the private LoRA weights indeed perform as claimed, without revealing them.

\section{Experiments}
We benchmarked \texttt{zklora} by measuring the creation and verification times for our verifiable computations across multiple base models and private LoRA configurations. For each model, we computed a forward pass for a batch of 3 items with sequence length 5. 
Each experiment simulates the workflow where the \emph{Base Model User} uses local data to evaluate the base model + LoRA combination, requesting remote LoRA computations from the \emph{LoRA Owner}. 
We report total and average proof-generation times (\texttt{total\_prove}, \texttt{avg\_prove}) and verification times (\texttt{total\_verify}, \texttt{avg\_verify}), as well as relevant model parameters.

\begin{table}[ht]
    \centering
    \small
    \begin{tabular}{l c c c c c c c}
    \toprule
    \textbf{base\_model} & \textbf{\#lora} & \textbf{total\_params} & \textbf{avg\_params} & \textbf{total\_prove} & \textbf{avg\_prove} & \textbf{total\_verify} & \textbf{avg\_verify} \\
    \midrule
distilgpt2 & 24 & 589824 & 24576.0 & 759.33 & 31.64 & 16.56 & 0.69 \\
gpt2 & 48 & 2359296 & 49152.0 & 1675.58 & 34.91 & 32.79 & 0.68 \\
Llama-3.2-1B & 32 & 851968 & 26624.0 & 991.93 & 31.00 & 24.91 & 0.78 \\
Llama-3.3-70B-Instr. & 80 & 11796480 & 147456.0 & 3749.76 & 46.87 & 123.11 & 1.54 \\
Llama-3.1-8B-Instr. & 32 & 5242880 & 163840.0 & 1527.40 & 47.73 & 35.79 & 1.12 \\
Mixtral-8x7B-Instr. & 32 & 10485760 & 327680.0 & 2357.61 & 73.68 & 44.30 & 1.38 \\
    \bottomrule
    \end{tabular}
    \caption{Benchmark results for \texttt{zklora} showing split inference overhead on various model + LoRA combinations. Times are in seconds.}
    \label{tab:results}
\end{table}

As shown in Table~\ref{tab:results}, proof-generation times tend to grow with model size, particularly for 70B-scale systems. However, the verification step remains fast, often averaging under two seconds per proof, even at large scales. This efficiency is crucial for real-world “trial runs”, allowing the Base Model User to confirm performance on a validation set without incurring prohibitive overhead. In larger-scale usage scenarios, once trust in the LoRA module is established, \texttt{zklora} may no longer be needed.

\section{Conclusion}
\texttt{zklora} enables a \emph{Base Model User} to verify the quality of a private LoRA module without revealing the LoRA weights. By pairing split inference with cryptographic proofs, \texttt{zklora} ensures that the user can trust the computed outputs during validation experiments. After verifying performance, the Base Model User may adopt any approach they choose to integrate or deploy the private LoRA more widely, since the module has been proven effective. Our experiments show this approach is feasible across a range of model sizes, with \emph{fast verification times} minimizing the overhead of a verifiable evaluation phase. Future directions include further circuit optimization and exploring multi-owner scenarios, in which multiple private LoRA modules could be composed and verified within the same pipeline.

>>>>

\title{\textbf{zklora: Verifiable Multi-Party Inference for Evaluating Private LoRA Weights}}
\author{Anonymous Authors}
\date{\today}

@@

\begin{abstract}
In decentralized computing environments, one party may wish to fine-tune a publicly available base model with Low-Rank Adaptation (LoRA) but keep those adapter weights private, while another party wants to evaluate performance on new data. 
We introduce \texttt{zklora}, a library designed to help the \emph{Base Model User} \emph{verifiably assess} the quality of a private LoRA module for a target task---for instance, to run a validation experiment without directly accessing the private weights. 
By adopting a multi-party inference approach and generating \emph{verifiable computations}, \texttt{zklora} enables the \emph{Base Model User} to confirm that remote LoRA computations are correct, all without seeing the private weights themselves. 
We demonstrate the feasibility of \texttt{zklora} via benchmark experiments, showing that our approach remains computationally viable across various model sizes, with notably fast verification times.
\end{abstract}

%------------------------------------------
% 1. PRELIMINARY RESULTS
%------------------------------------------
\section{Preliminary Results}

We benchmarked \texttt{zklora} by measuring the creation and verification times for our verifiable computations across multiple base models and private LoRA configurations. For each model, we computed a forward pass for a batch of 3 items with sequence length 5. 
Each experiment simulates the workflow where the \emph{Base Model User} uses local data to evaluate the base model + LoRA combination, requesting remote LoRA computations from the \emph{LoRA Owner}. 
We report total and average proof-generation times (\texttt{total\_prove}, \texttt{avg\_prove}) and verification times (\texttt{total\_verify}, \texttt{avg\_verify}), as well as relevant model parameters.

\begin{table}[ht]
    \centering
    \small
    \begin{tabular}{l c c c c c c c}
    \toprule
    \textbf{base\_model} & \textbf{\#lora} & \textbf{total\_params} & \textbf{avg\_params} & \textbf{total\_prove} & \textbf{avg\_prove} & \textbf{total\_verify} & \textbf{avg\_verify} \\
    \midrule
distilgpt2 & 24 & 589824 & 24576.0 & 759.33 & 31.64 & 16.56 & 0.69 \\
gpt2 & 48 & 2359296 & 49152.0 & 1675.58 & 34.91 & 32.79 & 0.68 \\
Llama-3.2-1B & 32 & 851968 & 26624.0 & 991.93 & 31.00 & 24.91 & 0.78 \\
Llama-3.3-70B-Instr. & 80 & 11796480 & 147456.0 & 3749.76 & 46.87 & 123.11 & 1.54 \\
Llama-3.1-8B-Instr. & 32 & 5242880 & 163840.0 & 1527.40 & 47.73 & 35.79 & 1.12 \\
Mixtral-8x7B-Instr. & 32 & 10485760 & 327680.0 & 2357.61 & 73.68 & 44.30 & 1.38 \\
    \bottomrule
    \end{tabular}
    \caption{Benchmark results for \texttt{zklora} showing multi-party inference overhead on various model + LoRA combinations. Times are in seconds.}
    \label{tab:results}
\end{table}

Notably, although proof-generation times can grow with model size, verification remains comparatively fast (e.g., \texttt{avg\_verify} of about 1--2 seconds even for 70B-scale models). This result demonstrates that \texttt{zklora} allows frequent correctness checks without excessive overhead, which is crucial for real-world “trial runs.” In larger-scale usage scenarios, once trust in the LoRA module is established, \texttt{zklora} may no longer be needed.

%------------------------------------------
% 2. MULTI-PARTY INFERENCE METHOD
%------------------------------------------
\section{Multi-Party Inference: zklora for Verifiable Evaluation}
Our aim is to enable the \emph{Base Model User} to run a \emph{validation experiment} on their local data, incorporating the private LoRA module from the \emph{LoRA Owner}. The Base Model User can measure task performance (e.g., cross-entropy loss, accuracy) to decide whether the LoRA meets performance requirements.

Figure~\ref{fig:toy-mlp-lora} illustrates a simple scenario:
\begin{enumerate}
    \item \textbf{Base Model Forward Pass (Base Model User).}
    The \emph{Base Model User} runs layers without LoRA parameters locally, generating partial activations.

@@

\draw[->] (input) -- node[above, font=\scriptsize]{local multiply} (W1);
\draw[->, dashed] (W1.north) to[bend left=15] node[left, font=\scriptsize]{partial activations} (LoRA.south);
\draw[->, dashed] (LoRA.south) to[bend left=15] node[right, font=\scriptsize]{updated $\mathbf{z}'$} (W1.north);
\draw[->] (W1) -- node[above, font=\scriptsize]{$\mathbf{z}'$} (act);
\draw[->] (act) -- node[above, font=\scriptsize]{hidden} (W2);
\draw[->] (W2) -- (output);

\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black]above:\textbf{Base Model User}},
      inner sep=0.4cm
] (baseModelFit) {};
\end{tikzpicture}
\caption{A simplified 2-layer MLP with LoRA on the first layer. The Base Model User (blue box) runs most computations locally but calls the LoRA Owner (dashed box) to update partial activations. After verifying correctness, the Base Model User can measure final metrics to decide whether to adopt the private LoRA more extensively.}
\label{fig:toy-mlp-lora}
\end{figure}

%------------------------------------------
% 3. INTRODUCTION
%------------------------------------------
\section{Introduction}
Large Language Models (LLMs) excel at a variety of tasks but come with significant computational and storage costs \cite{brown2020language, devlin2018bert}. 
Low-Rank Adaptation (LoRA) \cite{hu2021lora} addresses these costs by introducing a small number of trainable parameters into the model, drastically reducing memory overhead for fine-tuning \cite{ding2022delta}.

In this paper, we present \texttt{zklora}, a library that provides a \emph{verifiable computation} solution for the \emph{Base Model User} to \emph{evaluate} a private LoRA module. By orchestrating a multi-party inference, \texttt{zklora} allows the Base Model User to run the base model locally while sending partial activations to the LoRA Owner. Crucially, the LoRA Owner returns not only updated activations but also a cryptographic \emph{proof} that the correct LoRA transformation was applied. In this way, the Base Model User can confirm the final outputs truly reflect the LoRA module’s capabilities. After verifying and measuring results on a validation dataset, the Base Model User can later choose how to adopt the LoRA at larger scale.

%------------------------------------------
% 4. RELATED WORK
%------------------------------------------
\section{Related Work}

%------------------------------------------
% 5. CONCLUSION
%------------------------------------------
\section{Conclusion}
\texttt{zklora} enables a \emph{Base Model User} to verify the quality of a private LoRA module without revealing the LoRA weights. By pairing multi-party inference with cryptographic proofs, \texttt{zklora} ensures that the user can trust the computed outputs during validation experiments. After verifying performance, the Base Model User may adopt any approach they choose to integrate or deploy the private LoRA more widely, since the module has been proven effective. Our experiments show this approach is feasible across a range of model sizes, with \emph{fast verification times} minimizing the overhead of a verifiable evaluation phase. Future directions include further circuit optimization and exploring multi-owner scenarios, in which multiple private LoRA modules could be composed and verified within the same pipeline.
